---
title: "Survival Analysis"
author: "Jinfei Li"
date: "2021/12/8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(kableExtra.latex.load_packages = FALSE)
```

## K-M Curve and Simple Parametric Distribution

### Data Summary
Data can be downloaded from [here](https://www.kaggle.com/taimurzahid/survival-regression-analysis-to-predict-churn/data), in the dataset, there is no entry age, which is exist in real case, so I added entry_age = 0 for all observations. Entry age is the client's age (or loan age) when we start recording the records history.

```{r loaddata, echo=FALSE, include = FALSE, messsage = FALSE,  warning = FALSE}
library(data.table)
library(flexsurv)
library(ggplot2)
library(gridExtra)
library(eha)
library(msm)
library(CompLognormal)
library(pROC)
require(plyr)
library(kableExtra)
library(survival)
library(survminer)

setwd("C:/Users/luoyan011/Desktop/PersonalLearning/GitHub/data_2021")
data = fread('WA_Fn-UseC_-Telco-Customer-Churn.csv')
data$entry_age = 0


data$Churn <- mapvalues(data$Churn, 
          from=c('Yes','No'), 
          to=c(1,0))
data$Churn = as.numeric(as.character(data$Churn))

data = data[which(data$tenure>data$entry_age),]

index = sample(1:nrow(data), size=0.3*nrow(data))
training = data[-index,]
testing = data[index,]
```

```{r functions, echo=FALSE, include = FALSE, messsage = FALSE,  warning = FALSE}
logit.to.AIC = function(x){
  2-2*mean(x$loglik)
}

base.weib = function(x,t,i){
  # x is distibution
  # t is time index
  # i is for strata, 2 categories then i = 1 and 2, 3 categories then i = 1,2,3
  s1 = pweibull(t
                , shape = exp(x$coefficients[2*i])
                , scale = exp(x$coefficients[2*i-1])
                , lower.tail = FALSE
                )
  return(s1)
}

base.exp = function(x,t){
  exp(-exp(x$coefficients)*t)
}

base.lognom = function(x,t,i){
  s1 = plnorm(t
              , x$coefficients[2*i-1]
              , exp(-x$coefficients[2*i])
              , lower.tail = FALSE
              )
  return(s1)
}

base.loglog = function(x,t,i){
  s1 = pllogis(t
               , exp(x$coefficients[2*i])
               , exp(x$coefficients[2*i-1])
               , lower.tail = FALSE
               )
}


```

The dataset samples are as below, where tenure is age, Churn is target variable, entry_age is the control of entry-age, customerID is ID, all other variables are covariate variables.
```{r displaydata, echo=FALSE}
kbl(data[1:3,1:12], booktabs = T) %>%
  kable_styling(latex_options = c("striped", "scale_down"))
kbl(data[1:3,13:22], booktabs = T) %>%
  kable_styling(latex_options = c("striped", "scale_down"))

```

### Parametric Method -- Simple Distribution

Fit parametric distribution with no covariates:
```{r para_no_covariate, echo=TRUE, warning=FALSE}
exp.1 = flexsurvreg(Surv(entry_age, tenure, Churn, type='counting')~1
                    , data=training
                    , dist = 'exp')
lognormal.1 = aftreg(Surv(entry_age, tenure, Churn, type='counting')~1
                    , data=training
                    , dist = 'lognormal')
weibull.1 = aftreg(Surv(entry_age, tenure, Churn, type='counting')~1
                    , data=training
                    , dist = 'weibull')
loglog.1 = aftreg(Surv(entry_age, tenure, Churn, type='counting')~1
                    , data=training
                    , dist = 'loglogistic')
plot(lognormal.1)

```

Fit parametric distribution with variable "PhoneService":
```{r para_covariate, echo=TRUE, warning=FALSE}
lognormal.2 = aftreg(Surv(entry_age, tenure, Churn, type='counting')~strata(PhoneService)
                    , data=training
                    , dist = 'lognormal')
weibull.2 = aftreg(Surv(entry_age, tenure, Churn, type='counting')~strata(PhoneService)
                    , data=training
                    , dist = 'weibull')
loglog.2 = aftreg(Surv(entry_age, tenure, Churn, type='counting')~strata(PhoneService)
                    , data=training
                    , dist = 'loglogistic')
```

Check AIC for different models above:
```{r para_aic, echo=FALSE, warning=FALSE}
AIC.1 = cbind(logit.to.AIC(weibull.1),logit.to.AIC(lognormal.1),logit.to.AIC(loglog.1),logit.to.AIC(exp.1))
colnames(AIC.1) = c('weibreg.AIC','lognormal.AIC','loglogistic.AIC','exponential.AIC')
AIC.2 = cbind(logit.to.AIC(weibull.2),logit.to.AIC(lognormal.2),logit.to.AIC(loglog.2))
colnames(AIC.2) = c('weibreg.AIC','lognormal.AIC','loglogistic.AIC')
```

```{r display_paraaic, echo=FALSE}
AIC.1 %>%
  kbl(caption = "No Covariates Distribution AIC") %>%
  kable_classic(full_width = F, html_font = "Cambria")
AIC.2 %>%
  kbl(caption = "With Covariates Distribution AIC") %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

Get distribution parameters and some survival functions:
```{r para-parameters, echo=TRUE, warning=FALSE}
t = c(1:100)
# No-covariate-Weibull-distribution
shape = exp(weibull.1$coefficients[2])
scale = exp(weibull.1$coefficients[1])
s.weibull = exp(-((1/scale)*t)^shape)

# With-covariate-weibull-distrition
shape.1 = exp(weibull.2$coefficients[2])
scale.1 = exp(weibull.2$coefficients[1])
shape.2 = exp(weibull.2$coefficients[4])
scale.2 = exp(weibull.2$coefficients[3])
s.weibull.1 = exp(-((1/scale.1)*t)^shape.1)
s.weibull.2 = exp(-((1/scale.2)*t)^shape.2)

# No-covariate-lognormal-distribution
mu = lognormal.1$coefficients[1]
sigma = exp(lognormal.1$coefficients[2])

# No-covariates-loglog-distribution
alpha = exp(loglog.1$coefficients[1])
beta = exp(loglog.1$coefficients[2])
s.loglog = 1/(1+(t/alpha)^beta)

# Exponential-distribution
s.exp = exp(-exp(exp.1$coefficients)*t)
```

Get baseline survival rate and generate plot:
```{r para=plot, echo=FALSE, warning=FALSE}
base.exp.1 = base.exp(exp.1,t)

base.weibull.1 = base.weib(weibull.1, t, 1)
base.weibull.2.1 = base.weib(weibull.2, t, 1)
base.weibull.2.2 = base.weib(weibull.2, t, 2)

base.lognormal.1 = base.lognom(lognormal.1, t, 1)
base.lognormal.2.1 = base.lognom(lognormal.2, t, 1)
base.lognormal.2.2 = base.lognom(lognormal.2, t, 2)

base.logistic.1 = base.loglog(loglog.1, t, 1)
base.logistic.2.1 = base.loglog(loglog.2, t, 1)
base.logistic.2.2 = base.loglog(loglog.2, t, 2)

compare.base.1 = as.data.frame(cbind(t, base.weibull.1, base.exp.1, base.logistic.1, base.lognormal.1))
ggplot(compare.base.1
       , aes(t)) +
  geom_line(aes(y = base.weibull.1, colour = 'weibull')) +
  geom_line(aes(y = base.exp.1, colour = 'exp')) +
  geom_line(aes(y = base.logistic.1, colour = 'loglog')) +
  geom_line(aes(y = base.lognormal.1, colour = 'lognormal')) +
  labs(title = 'Baseline Survival Function without Covariates', x = 'Time',y='Survival Rate')

compare.base.2 = as.data.frame(cbind(t, 
                                     base.weibull.2.1, base.weibull.2.2,
                                     base.logistic.2.1, base.logistic.2.2,
                                     base.lognormal.2.1, base.lognormal.2.2))
ggplot(compare.base.2, aes(t)) +
  geom_line(aes(y = base.weibull.2.1, colour = 'weibull', linetype = 'No-phone-service')) +
  geom_line(aes(y = base.weibull.2.2, colour = 'weibull', linetype = 'Phone-service')) +
  geom_line(aes(y = base.logistic.2.1, colour = 'loglog', linetype = 'No-phone-service')) +
  geom_line(aes(y = base.logistic.2.2, colour = 'loglog', linetype = 'Phone-service')) +
  geom_line(aes(y = base.lognormal.2.1, colour = 'lognormal', linetype = 'No-phone-service')) +
  geom_line(aes(y = base.lognormal.2.2, colour = 'lognormal', linetype = 'Phone-service')) +
  labs(title = 'Baseline Survival Function with Covariates', x = 'Time',y='Survival Rate')
```

Check assumption of distribution
```{r para-assumption, echo=FALSE, warning=FALSE}
ggplot(compare.base.1, aes(log(t))) +
  geom_line(aes(y=log(-log(base.weibull.1)), colour = 'weibull')) +
  geom_line(aes(y=log(-log(base.logistic.1)), colour = 'loglog')) +
  geom_line(aes(y=log(-log(base.lognormal.1)), colour = 'lognormal')) +
  labs(title = 'Assumption for distribution with no-covariates', x = 'Log Time',y='Log Survival Rate')
```

### KM Curve / CoxPH Model

Cox-PH model code example can be found from [here](http://www.sthda.com/english/wiki/cox-proportional-hazards-model)

```{r kmfit, echo=TRUE, warning=FALSE}
base.coxph = coxph(Surv(entry_age, tenure, Churn) ~ 1, data = training)
base.hazard = basehaz(base.coxph)
base.km = exp(-base.hazard$hazard)
```